Joshua 2.0: A Toolkit for Parsing-Based Machine Translation \\with Syntax, Semirings, Discriminative Training and Other Goodies

### 2010-04-23 papers/wmt10/joshua-progress.tex
Revising Zhifei's text
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1631 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

\textbf{Joshua} is an open-source toolkit for parsing-based machine translation that is written in Java. The initial release of Joshua  \cite{joshua-wmt09} was a re-implementation of the Hiero system \cite{hiero-2007} and all its associated algorithms, including: chart-parsing, $n$-gram language model integration, beam- and cube-pruning, and $k$-best extraction.  The Joshua 1.0 release also included re-implementations of suffix-array grammar extraction \cite{lopez:2007:EMNLP-CoNLL2007,PBML-2010-Joshua-grammar-extraction}  and minimum error rate training \cite{och-mert,Zaidan2009}.  Additionally, it included parallel and distributed computing techniques for salability \cite{Joshua-old}.

This paper describes the additions to the toolkit over the past year, which together form the 2.0 release. The software has been heavily used by the authors and several other groups in their daily research, and has been substantially refined since the first release.  The most important new functions in the toolkit are: 
\begin{itemize}
\item
Support for any style of synchronous context free grammar (SCFG)
% instead of just Hiero grammars.  This 
including syntax augment machine translation (SAMT) grammars \cite{samt2006}
\item
Support for external modules to posit translations for spans in the input sentence that constrain decoding   \cite{PBML-2010-Joshua-transliteration}
\item
Lattice parsing for dealing with input uncertainty, including ambiguous output from speech recognizers or Chinese word segmenters  \cite{dyer-muresan-resnik:2008:ACLMain}
\item
A semiring architecture over hypergraphs that allows many inference operations to be implemented easily and elegantly \cite{li-eisner:2009:EMNLP}
\item
Improvements to decoding through variational decoding and other approximate methods that overcome intractable MAP decoding \cite{variational-decoding-acl09}
\item
Hypergraph-based discriminative training for better feature engineering \cite{zhifei-forest-reranking-galebook}
\item
Visualization of the derivation trees and hypergraphs \cite{PBML-2010-Joshua-visualization}
\item
A convenient framework for designing and running reproducible machine translation experiments \cite{Schwartz-wmt10-pipline}
\end{itemize}
The sections below give a short description of each of these new functions.

@@

Integrating output from specialized modules (like transliterators, morphological analyzers, and modality translators) into the MT pipeline can improve translation performance, particularly in low-resource settings. We have implemented this functionality in Joshua by supporting an XML format for proposing alternate translation rules (constraints) for a particular word span to the decoder \cite{pbml-2010-Joshua-modules}. That is, processing that is separate from the MT engine can suggest translations for some set of source side words and phrases. The XML format allows for both hard constraints, which must be used, and soft constraints, which compete with standard extracted translation rules, as well as specifying associated feature weights. We modified the chart-based decoder to support these constraints.

@@

In \joshua, we use a {\em hypergraph} (or {\em packed forest}) to compactly 
represent the exponentially many derivation trees generated by the decoder for a 
particular foreign input sentence.
Given a hypergraph, we may perform many atomic inference operations
such as finding one-best, $k$-best, or expectations over the hypergraph.
To perform each operation, we may implement a dedicated dynamic programming algorithm.
However, a more general framework to specify these algorithms is semiring-weighted parsing \cite{semiringparsing}.
Within this framework, we implemented all the algorithms (i.e.,
inside, outside, and inside-outside speedup) described by \newcite{li-eisner:2009:EMNLP},
and also the first-order expectation semiring \cite{eisner-expectation-semiring} and its
second-order version \cite{li-eisner:2009:EMNLP}.

@@

%% use of expectation and variance semiring
The first- and second-order expectation semirings can be used compute many interesting quantities over the hypergraph.
These quantities include expected hypothesis length, feature expectation, entropy, cross-entropy, Kullback-Leibler divergence,
Bayes risk, variance of hypothesis length, gradient of entropy and Bayes risk, covariance and Hessian matrix, and so on.

@@

We have introduced tools for visualizing two of the main data structures used
in Joshua \cite{PBML-2010-Joshua-transliteration}. Each visualization tool is implemented as a
stand-alone program that hooks into the pipeline, but they are distributed with
Joshua. The first program is a hypergraph visualizer. The user can choose from
a set of input sentences, then call the decoder to build the hypergraph.

>>>>

\textbf{Joshua} is an open-source toolkit for parsing-based machine translation that is written in Java. The initial release of Joshua  \cite{joshua-wmt09} was a re-implementation of the Hiero system \cite{hiero-2007} and all its associated algorithms, including: chart parsing, $n$-gram language model integration, beam and cube pruning, and $k$-best extraction.  The Joshua 1.0 release also included re-implementations of suffix array grammar extraction \cite{lopez:2007:EMNLP-CoNLL2007,PBML-2010-Joshua-grammar-extraction}  and minimum error rate training \cite{och-mert,Zaidan2009}.  Additionally, it included parallel and distributed computing techniques for salability \cite{Joshua-old}.

This paper describes the additions to the toolkit over the past year, which together form the 2.0 release. The software has been heavily used by the authors and several other groups in their daily research, and has been substantially refined since the first release.  The most important new functions in the toolkit are: 
\begin{itemize}
\item
Support for any style of synchronous context free grammar (SCFG)
% instead of just Hiero grammars.  This 
including syntax augment machine translation (SAMT) grammars \cite{samt2006}
\item
Support for external modules to posit translations for spans in the input sentence that constrain decoding   \cite{PBML-2010-Joshua-modules}
\item
Lattice parsing for dealing with input uncertainty, including ambiguous output from speech recognizers or Chinese word segmenters  \cite{dyer-muresan-resnik:2008:ACLMain}
\item
A semiring architecture over hypergraphs that allows many inference operations to be implemented easily and elegantly \cite{li-eisner:2009:EMNLP}
\item
Improvements to decoding through variational decoding and other approximate methods that overcome intractable MAP decoding \cite{variational-decoding-acl09}
\item
Hypergraph-based discriminative training for better feature engineering \cite{zhifei-forest-reranking-galebook}
\item
Visualization of the derivation trees and hypergraphs \cite{PBML-2010-Joshua-visualization}
\item
A convenient framework for designing and running reproducible machine translation experiments \cite{Schwartz-wmt10-pipline}
\end{itemize}
The sections below give short descriptions for each of these new functions.

@@

Integrating output from specialized modules (like transliterators, morphological analyzers, and modality translators) into the MT pipeline can improve translation performance, particularly in low-resource settings. We have implemented this functionality in Joshua by supporting an XML format for proposing alternate translation rules (constraints) for a particular word span to the decoder \cite{PBML-2010-Joshua-modules}. That is, processing that is separate from the MT engine can suggest translations for some set of source side words and phrases. The XML format allows for both hard constraints, which must be used, and soft constraints, which compete with standard extracted translation rules, as well as specifying associated feature weights. We modified the chart-based decoder to support these constraints.

@@

In \joshua, we use a {\em hypergraph} (or {\em packed forest}) to compactly 
represent the exponentially many derivation trees generated by the decoder for a 
particular foreign input sentence.
Given a hypergraph, we may perform many atomic inference operations,
such as finding one-best or $k$-best translations, or computing expectations over the hypergraph.
For each such operation, we could implement a dedicated dynamic programming algorithm.
However, a more general framework to specify these algorithms is semiring-weighted parsing \cite{semiringparsing}.
We have implemented  the inside algorithm, the outside algorithm, and the inside-outside speedup described by \newcite{li-eisner:2009:EMNLP},
plut the first-order expectation semiring \cite{eisner-expectation-semiring} and its
second-order version \cite{li-eisner:2009:EMNLP}.  All of these use our newly implemented semiring framework.

@@

%% use of expectation and variance semiring
The first- and second-order expectation semi-rings can also be used to compute many interesting quantities over hypergraphs.
These quantities include expected translation length, feature expectation, entropy, cross-entropy, Kullback-Leibler divergence,
Bayes risk, variance of hypothesis length, gradient of entropy and Bayes risk, covariance and Hessian matrix, and so on.

@@

We have introduced tools for visualizing two of the main data structures used
in Joshua \cite{PBML-2010-Joshua-visualization}. Each visualization tool is implemented as a
stand-alone program that hooks into the pipeline, but they are distributed with
Joshua. The first program is a hypergraph visualizer. The user can choose from
a set of input sentences, then call the decoder to build the hypergraph.
